---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)
library(tidyverse)
```

```{css}
h2 {
  font-weight: bold;
  text-align: center;
  color: #9FD5E8
}

h3 {
  color: #9FD5E8
}

p{
  padding: 30pt
}

body{
  font-family: Sans-Serif;
  font-size: 14pt;
  padding: 20pt;
}

```

## Bird Diversity at Meola Reef Reserve (Te Tokaroa)

This study was done on the abundance and distribution of native and non-native bird species at Meola Reef Reserve (Te Tokaroa) in Western Springs, during March 2025. The purpose of this was to observe and understand more about the diversity of bird species at Meola Reef Reserve, a suburban location in Auckland, New Zealand. This study can have future benefits as this data can be used to create favourable habitats to support a higher diversity of native species.

```{r}
logged_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSEcG28d-pAJ7jbyBIkgvRrlLHbVzawzAD0k6aOz-vRThTM_iI-umNRR5KBcFqHDxKRP8MOCw2cY1pd/pub?output=csv")

latest_data <- logged_data %>%
  rename(NZ_status = 2,
         frequency = 3,
         habitat = 4)
```

### Summary Values

`r paste("The total number of birds observed during this study was", latest_data$frequency %>% sum(), "individuals.")` \
`r paste("The average number of birds of one species observed during this study was", latest_data$frequency %>% mean() %>% round(2), "individuals.")`

### Abundance Graph

The first factor analysed was abundance. This bar graph compares the number of native birds to non-native birds at Meola Reef Reserve.

```{r abundance, fig.width = 13.5, fig.height = 7, fig.align = "center"}
ggplot(data = latest_data) +
  geom_bar(aes(x = NZ_status),
           fill = c("#9FD5E8", "#A1BD8B")) +
  labs(title = "Abundance of native and non-native bird species found at Meola Reef Reserve",
       subtitle = "Observed during March 2025",
       x = "New Zealand status",
       y = "Number of birds") +
  theme(
    plot.title = element_text(size = 20, face = "bold"),
    plot.subtitle = element_text(size = 16),
    axis.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 14)
  )
```

### Distribution Graph

The second factor analysed was distribution. This bar graph shows the habitat distribution of birds found at Meola Reef Reserve, and whether they were native or non-native species.

```{r distribution, fig.width = 13.5, fig.height = 7, fig.align = "center"}
ggplot(data = latest_data) +
  geom_bar(aes(x = habitat, fill = NZ_status)) +
  labs(title = str_wrap("Habitat distribution of bird species found at Meola Reef Reserve, grouped by their New Zealand status",
                        width = 90),
       subtitle = "Observed during March 2025",
       x = "Habitat (vegetation area type)",
       y = "Number of birds",
       fill = "New Zealand Status") +
  scale_fill_manual(values = c("#9FD5E8", "#A1BD8B")) +
  theme(
    plot.title = element_text(size = 20, face = "bold"),
    plot.subtitle = element_text(size = 16),
    axis.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 14)
  ) +
  scale_x_discrete(
    labels = c("Native", "Rank grass", "Salt marsh, coastal shrub or mangrove"))
```
